How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985

### 2022-04-14 Draft_April_2022/Full_draft_20220401.rmd
Corrected a few footnotes, added note for regression tibble, and re-added sections 4.3 and conclusion

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Both of them are PhD candidates at The Geneva Graduate Institute. Livio is affiliated to the Albert Hirschman Centre on Democracy. Henrique is affiliated to the Centre for International Environmental Studies.]
date: "April 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: This paper ...
---

@@

Problem-construction at the level of discourse is also more varied. They are not monolithic in time, across location, or even by the same speaker. While governmental discourses in Brazil have been studied for topic such as inflation or race relations [@hunt2004, @silva2019, and @djik2020], we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

@@

kbl(codebook, booktabs = T,caption = "Table 1 - Amazonian Problem-Construction Codebook")%>% kable_styling(latex_options = c("striped", "repeat_header", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="3cm", bold=TRUE)%>%
  column_spec(2, width="9cm") %>%
  column_spec(3, width = "9cm")

@@

We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories [@meyer2021; @noble2006][^3]. After the SVM model is trained and tuned, and the model shows to appropriately fit the validation set, we use the model to automatically code the remaining 1007 Amazonian statements. The model was also trained to identify false positive observations, which we chose to delete for analysis since these are not meaningful statements about the Amazon. The final dataset for analysis contains 1895 coded Amazonian statements. Finally, we use poldis once more to extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

[^3]: For more information on this please see the extended methodology. The document includes a detailed discussion of how the SVM algorithm was chosen over other algorithms based on evaluation of performance of trained models for labeling the validation set. The extended methodology also includes a more detailed codebook table with further coding guidelines.

@@

Figure 1 shows the proportion of speeches that mentions the Amazon in relation to all speeches in each year. The predicted share curve controls incidence for deforestation[^4] , economic situation, and speaker (see appendix for methodological details). We observe various local maxima: 1989, 1992, 2005, 2009, 2015, and 2019. These points coincide with events that help us explain the rises and falls of the Amazon in presidential discourse.

[^4]: The reasoning here is that knowledge of high or low deforestation rates might drive presidents to speak about the Amazon. Nevertheless, matching deforestation with speech dates is a complicated endeavor. INPE published consolidated deforestation data with almost two years of delay until 2005 [@capobianco2021, pp. 60]. Preliminary data was published earlier, but with lower degrees of confidence in the findings. For the subsequent years, the consolidated figures for a given year came out in August of the subsequent year. However, it also seems to be the case that the executive government has access to the data before everyone else. In addition, other sources indicating if deforestation is going up or down, such as fire data or lower resolution deforestation data (DETER for example), or even other sources, circulated within the year the data covered. The year of 1988 is particularly indicative of these complications: a report dating 1988 was circulating with deforestation and fire figures [@fearnside1990], and a New York Times article about the issue with deforestation rates for 1988 was written with comments by an INPE scientist [@simons1988]. Hence, we believe the most appropriate way of matching is without any lags on deforestation.

Figure 1 shows the proportion of speeches that mentions the Amazon in relation to all speeches in each year. The predicted share curve controls incidence for deforestation[\^4], economic situation, and speaker (see appendix for methodological details). We observe various local maxima: 1989, 1992, 2005, 2009, 2015, and 2019. These points coincide with exogenous events that helps us explain the rises and falls of the Amazon in presidential discourse.

@@

@@

Figure 5, above, illustrates the share of pure-type problem construction by locations in Brazil. We divide locations into Amazonian states (i.e. all the Brazilian states in which the Amazon biome is present), non-Amazonian states (other brazilian states), Brasilia (the capital city in Brazil), and international (countries outside of Brazil). Notice that, consistently, presidents construct the Amazon as an issue of pure economic integration within Amazonian states more often than all other constructions in time. Alternatively, we see that pure conservation becomes increasingly less frequent in time for Amazonian states. As well, pure social development constructions, which were fairly frequent from the early-2000s and onwards, disappeared around 2015 after the mid-2010s in Amazonian states. Nonetheless, we see a large increase from the mid-2000s to the mid-2010s in the frequency in which presidents constructed the Amazon as an issue of pure social development in Brasilia and non-Amazonian states. In Brasilia, also, pure conservation constructions went down drastically during the 1990s, but increased steadily from the mid-2000s onwards. Interestingly, presidents most often constructed the Amazon as an issue of pure environmental conservation internationally only from the mid 2000s until the mid-2010s. Moreover, pure sovereignty constructions did appear in international settings until the mid-2000s.

Table 2, below, displays the multinomial regression coefficients for the correlation between diverse problem constructions and locations. The regression coefficients express the log odds, that is, the logarithm odds to the reference categories. The reference categories are Amazonian states for location and environmental conservation for problem construction. To facilitate visualization, since the model has many more columns (i.e. types of problem construction), we only display the model results for the pure-type constructions, however, the model does account for both pure and mixed-type constructions. We keep the results for the control variables for election year, annual deforestation rates, and average yearly inflation in the table.

@@

kbl(t(md), booktabs = T,caption = "Table 2 - Amazon Problem-Construction by Location", align = "c") %>%
  kable_styling(latex_options = c("striped", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="5 cm", bold=TRUE, latex_valign = "m")%>%
  column_spec(2, width="5cm", latex_valign = "m") %>%
  column_spec(3, width = "5cm", latex_valign = "m") %>%
  column_spec(4, width = "5cm", latex_valign = "m")
```

The results indicate that the presidents are more likely to construct it as an environmental conservation problem far away from the Amazon. The negative statistically significant coefficients of problem constructions such as pure economic integration and pure social development in both international settings and Brasilia, in relation to environmental conservation problem constructions in Amazonian states, indicate that presidents are less likely to construct the Amazon as an issue of economic integration and social development in those locations. That is, presidents are more likely to construct the Amazon as an issue of environmental conservation in international settings and in Brasilia than in Amazonian states.

```{r Figure 6: probability for constructions across locations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}

logit2prob <- function(logit) {
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}
md_plot <- data.frame("Construction" = c("Economic Integration", "Social Development", "Sovereignty"),
                      "Brasilia" = c(-1.024, -1.002, -0.206),
                      "International" = c(-0.612, -1.583, -1.427),
                      "Other_States" = c(0.023, -0.041, 0.324))
md_plot$Brasilia <- logit2prob(md_plot$Brasilia)
md_plot$International <- logit2prob(md_plot$International)
md_plot$Other_States <- logit2prob(md_plot$Other_States)
mplot <- md_plot %>% tidyr::pivot_longer(Brasilia:Other_States) %>%
  rename(Location = name, Probability = value) %>% 
  ggplot(aes(Construction, Probability)) +
  geom_bar(aes(fill = Location), stat = "identity", position = position_dodge(0.8), width = 0.7) +
    scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "Probability",
       title = "Probability for problem constructions across locations",
       subtitle = "Reference category is environmental conservation problem-construction in Amazonian States") +
    theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
mplot
```

Sovereignty in the Amazon is an issue within Brazil. The negative statistically significant coefficient for constructing the Amazon as a problem of pure sovereignty in international settings, in relation to conservation problem constructions in Amazonian states, indicate that presidents are less likely to construct the Amazon as an issue of economic integration and social development in those locations. The lack of statistically significant correlations between pure sovereignty constructions and Brasilia or other non-Amazonian states, corroborates that the Amazon as an issue of sovereignty is an internally performed issue.

Constructing the Amazon as an issue of economic integration correlates with increases in deforestation rates. Some of the control variables correlate with some Amazonian problem constructions in interesting ways. For example, pure sovereignty constructions correlate negatively with election years in relation to conservation. That is, in election years conservation constructions are more likely to take place in comparison to sovereignty. As well, annual deforestation rates correlate positively with economic integration constructions in relation to conservation. This indicates that as deforestation increases, presidents are also more likely to construct the Amazon as an issue of economic integration, or vice-versa.

@putnam1988 seminal article on the two level game between domestic and international politics analyzes how negotiations at both levels make policies possible. The author argues that both domestic and international levels need to be taken into consideration when analyzing how, often, domestic politics became entangled in international negotiations. The Amazon, as a region and a forest, has been the topic of international negotiations, national debates, and local policy implementation; though how the Amazon has been constructed as an issue differs on each of these three levels. While in Brasilia or internationally different presidents might construct the Amazon as an issue of conservation, the same presidents might construct the Amazon as an issue of economic integration or social development at the local level. This implies not only that audiences' priorities in each setting change, but that which policies are appropriate to solve the "Amazon issue" differ. The three level game entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not for local electorates. Such might help explain diverse local implementation gaps between public policy negotiated outside and implemented within the Amazon as perceptions and expectations about the issues the same policy addresses might differ [@alesina2009; @lopez2020].

Whereas presidential discourses at the top matter to define and justify public policy [@zarefsky2004], presidents shape their discourses according to who their audience might be and what "they want to hear". This democratic game contributes to the advancement of ambiguous public policies thought from the outside to the Amazon or contradictory policies from Amazon to the outside. Take, for example, the rural credit offered to local agricultural producers in Amazonian states went from 500 million reais in 1999 to over 4 billion by 2012 [@capobianco2021]. During the same period, the money spent in fighting deforestation in the Amazonian states also increased from ... in ... to ... in ... These policies match diverse local, national, and international expectations and match solutions to different problem constructions presented at each of these levels.

@@

@@

stargazer::stargazer(model, type = "text",
                     covariate.labels = c("Brasilia", "International", "Non-AM state",
                                          "Election", "Deforestation", "Inflation"),
                     column.labels = c("EI", "SD", "Sov", "EI-Con", "EI-SD", "Sov-Con", "SD-EI-Con",
                                       "Sov-EI", "SD-Con", "Other"),
                     dep.var.labels.include = FALSE)

@@

  kable_styling(latex_options = c("striped", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="5 cm", bold=TRUE, latex_valign = "m")%>%
  column_spec(2, width="5cm", latex_valign = "m") %>%
  column_spec(3, width = "5cm", latex_valign = "m") %>%
  column_spec(4, width = "5cm", latex_valign = "m")

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? The economy, the environment, the people, and the nation in the presidential speeches since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article; names are ordered alphabetically. Both of them are PhD candidates at The Geneva Graduate Institute. Livio is affiliated to the Albert Hirschman Centre on Democracy. Henrique is affiliated to the Centre for International Environmental Studies.]
date: "April 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document:
    reference_docx: template.docx
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
#abstract: This paper ...
---

@@

Problem-construction at the level of discourse is also more varied. They are not monolithic in time, across location, or even by the same speaker. While governmental discourses in Brazil have been studied for topic such as inflation [@porto2007, @malheiros1995, and @fonseca2021] or race relations [@htun2004, @silva2019, and @djik2020], we only find one systematic analysis of Amazonian discourse. @barros2020 investigates Amazonian discourse in the Brazilian Congress with the objective of identifying the arguments put forth by congressmen. The main finding is that the economic value of the Amazon for the cattle industry is the most salient narrative, leading the author to conclude there is a mismatch between the international debate (which focuses on preservation) and the national debate (which focuses on economic development).

@@

kbl(codebook, booktabs = T, caption = "Table 1 - Amazonian Problem-Construction Codebook") %>% kable_styling(latex_options = c("striped", "repeat_header", "scale_down"),position= "center", font_size = 9, full_width = TRUE) %>% column_spec(1, width="3cm", bold=TRUE) %>% column_spec(2, width="9cm") %>% column_spec(3, width = "9cm")

@@

We chose to employ a support-vector machine (SVM) algorithm to label texts, that is, a non-probabilistic linear classifier that classifies documents by assigning points in mapped space to maximize the gap between binary categories [@meyer2021; @noble2006][^2]. After the SVM model is trained and tuned, and the model shows to appropriately fit the validation set, we use the model to automatically code the remaining 1007 Amazonian statements. The model was also trained to identify false positive observations, which we chose to delete for analysis since these are not meaningful statements about the Amazon. The final dataset for analysis contains 1895 coded Amazonian statements. Finally, we use poldis once more to extract locations for all speeches in the data. These locations represent the Brazilian state in which certain speech was given or an international country.

[^2]: For more information on this please see the extended methodology. The document includes a detailed discussion of how the SVM algorithm and the performance of trained models for labeling the validation set. The extended methodology also includes a more detailed codebook table with further coding guidelines.

@@

Figure 1 shows the proportion of speeches that mentions the Amazon in relation to all speeches in each year. The predicted share curve controls incidence for deforestation[^3] , economic situation, and speaker (see appendix for methodological details). We observe various local maxima: 1989, 1992, 2005, 2009, 2015, and 2019. These points coincide with events that help us explain the rises and falls of the Amazon in presidential discourse.

[^3]: The reasoning here is that knowledge of high or low deforestation rates might drive presidents to speak about the Amazon. Nevertheless, matching deforestation with speech dates is a complicated endeavor. INPE published consolidated deforestation data with almost two years of delay until 2005 [@capobianco2021, pp. 60]. Preliminary data was published earlier, but with lower degrees of confidence in the findings. For the subsequent years, the consolidated figures for a given year came out in August of the subsequent year. However, it also seems to be the case that the executive government has access to the data before everyone else. In addition, other sources indicating if deforestation is going up or down, such as fire data or lower resolution deforestation data (DETER for example), or even other sources, circulated within the year the data covered. The year of 1988 is particularly indicative of these complications: a report dating 1988 was circulating with deforestation and fire figures [@fearnside1990], and a New York Times article about the issue with deforestation rates for 1988 was written with comments by an INPE scientist [@simons1988]. Hence, we believe the most appropriate way of matching is without any lags on deforestation.

Figure 1 shows the proportion of speeches that mentions the Amazon in relation to all speeches in each year. The predicted share curve controls incidence for deforestation, economic situation, and speaker (see appendix for methodological details). We observe various local maxima: 1989, 1992, 2005, 2009, 2015, and 2019. These points coincide with exogenous events that helps us explain the rises and falls of the Amazon in presidential discourse.

@@

Figure 5, below, illustrates the share of pure-type problem construction by locations in Brazil. We divide locations into Amazonian states (i.e. all the Brazilian states in which the Amazon biome is present), non-Amazonian states (other Brazilian states), Brasilia (the capital of Brazil where the federal government is located), and international (countries outside of Brazil). The plots reveal several trends. Presidents consistently construct the Amazon as an issue of pure economic integration within Amazonian states more often than all other constructions in time. Relatedly, we observe a steady decrease since the mid-1980s in pure conservation constructions in Amazonian states. This differentiates Amazonian states from national or international settings, where the relationship among problem-constructions appears less hierarchical. In Brasilia, for once, the conservation problem construction remains the most frequent since the mid 2000s. Outside of Brazil, conservation was the most used construction from the mid-2000s to the mid-2010s. Nonetheless, pure social development constructions, which increased in the early-2000s to become the second most used construction within Amazonian states, disappeared around 2015. In line with this, we also see a large increase in the frequency of social development constructions in Brasilia and non-Amazonian states from the mid-2000s to the mid-2010s, but not internationally. This implies that constructing the Amazon as an issue of social development is a construction deployed within the country. Alternatively, while there is some variation in the frequency in which pure sovereignty constructions were employed in Amazonian States, other States, and Brasilia, this problem-construction has been increasing since the early 2010s in all these settings. In international settings, however, pure sovereignty did not appear until the mid-2000s. As with social development, sovereignty seems to be employed domestically as a problem-construction for the Amazon.

@@

To ensure these results are robust, we run a multinomial regression. Our dependent variable is problem constructions (categorical), while our main dependent variable is location (categorical). The reference category for the multinomial model is environmental conservation in Amazonian states. As in section 4.1, we add controls for deforestation rates (numerical), inflation (numerical), and election year. Table 2 displays our results in log odds. Since the model accounts for both mixed and pure-type Amazonian problem constructions, and this makes for a large table, we only display the results for pure-types to facilitate visualization [^4]. Figure 6 shows the relative probability of problem-constructions and location in relation to environmental conservation in Amazonian states. For example, the bar for economic integration in other states represents a 50% percent probability of using this problem-construction in relation to our reference category. Hence, the lower the probabilities displayed in bars in figure 6, the less likely presidents are to construct the Amazon as an issue of environmental conservation in Amazonian states.

[^4]: Please find the full regression table in the appendix below.

@@

kbl(t(md), booktabs = T, caption = "Table 2 - Amazon Problem-Construction by Location", align = "c") %>%
  kable_styling(latex_options = c("striped", "scale_down"), position= "center", font_size = 9, full_width = TRUE) %>%
  footnote(general = "The multinomial regression coefficients (in log odds) and standard errors (in parentheses) are displayed. The asterisk marks the level of statistical significance for coefficient, multiple asterisks denote a statistically significant P value above the 95th percentile.") %>% 
  column_spec(1, width="5 cm", bold=TRUE, latex_valign = "m") %>%
  column_spec(2, width="5cm", latex_valign = "m") %>%
  column_spec(3, width = "5cm", latex_valign = "m") %>%
  column_spec(4, width = "5cm", latex_valign = "m")
```

```{r Figure 6: probability for constructions across locations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
logit2prob <- function(logit) {
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}
md_plot <- data.frame("Construction" = c("Economic Integration", "Social Development", "Sovereignty"),
                      "Brasilia" = c(-1.024, -1.002, -0.206),
                      "International" = c(-0.612, -1.583, -1.427),
                      "Other_States" = c(0.023, -0.041, 0.324))
md_plot$Brasilia <- logit2prob(md_plot$Brasilia)
md_plot$International <- logit2prob(md_plot$International)
md_plot$Other_States <- logit2prob(md_plot$Other_States)
mplot <- md_plot %>% tidyr::pivot_longer(Brasilia:Other_States) %>%
  rename(Location = name, Probability = value) %>% 
  ggplot(aes(Construction, Probability)) +
  geom_bar(aes(fill = Location), stat = "identity", position = position_dodge(0.8), width = 0.7) +
    scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "Probability",
       title = "Probability for problem constructions across locations",
       subtitle = "Reference category is environmental conservation problem-construction in Amazonian States") +
    theme(panel.background = element_rect("white", "black", .5, "solid"),
        panel.grid.major = element_line(color = "grey", size = 0.3, linetype = "solid"),
        axis.text = element_text(color = "black", size = 10),
        title = element_text(color = "black", size = 10, face = "bold"),
        legend.title = element_blank(),
        plot.subtitle = element_text(color = "black", size = 9, face = "plain"),
        legend.position = "bottom")
mplot
```

The results of our model corroborate some of the visible patterns in figure 5. Presidents are more likely to construct the Amazon as an issue of environmental conservation when they are far away from the Amazon. The negative statistically significant coefficients of problem constructions such as pure economic integration and pure social development in both international settings and Brasilia, in relation to environmental conservation problem constructions in Amazonian states, indicate that presidents are less likely to construct the Amazon as an issue of economic integration and social development in those locations. Figure 6 shows that in international settings, constructions of social development are 33% less likely than in Amazonian states. As well, we see that economic integration constructions are 15% less likely in international settings. In Brasilia, both economic integration and social development constructions are 13% less likely than in Amazonian states.

Constructing the Amazon as an issue of sovereignty happens within Brazil. The negative statistically significant coefficient (table 2) for constructing the Amazon as a problem of pure sovereignty in international settings, in relation to Amazonian states, indicates that presidents are less likely to construct the Amazon as an issue of economic integration and social development in those locations. Figure 6 illustrates that presidents are 21% less likely to construct the Amazon as an issue of sovereignty internationally in relation to Amazonian states. The probabilities of sovereignty constructions around 0.5 (equal probability) for Brasilia and other non-Amazonian states corroborates that the Amazon as an issue of sovereignty is an internally performed issue.

Nevertheless, we see that some of the control variables (table 2) correlate with some Amazonian problem constructions in interesting ways. For example, pure sovereignty constructions correlate negatively with election years in relation to conservation. Also, annual deforestation rates also correlate positively with economic integration constructions in relation to conservation. This indicates that the Amazon is more likely to be constructed as an issue of economic integration as deforestation increases.

The Amazon, as a region and a forest, has been the topic of international negotiations, national debates, and local policy implementation; though our results show that how the Amazon has been constructed as an issue differs on each of these three levels. @putnam1988 seminal article on the two-level game between domestic and international politics analyzes how negotiations at both levels make policies possible. The author argues that both domestic and international levels should be considered. Agreements are made in the overlap between the position of domestic actors and international actors; so negotiators play a two-level game to maximize the area of overlap, bringing both to a more moderate position in issues [@putnam1988]. This implies not only that audiences’ priorities in each setting change, but that which policies are appropriate to solve the “Amazon issue” could differ as well. The three level game entails that conservation might be a desirable construction when speaking internationally about the Amazon, or in Brasilia, but not for local electorates.

This three-level game led to different policies being negotiated and maximized at the interplay between the levels. Take, for once, how the Brazilian government successfully funded domestic public policy with international support since the 1990s (e.g. 1992 Programa Piloto para Proteção das Florestas Tropicais), by creating a federal fund for the protected areas (e.g. 2001 Amazon Region Protected Areas program), and by assuring international funds with the establishment of the Amazon Fund in 2008 which provided over USD1 billion for environmental conservation in the Amazon [@silva-muller2022]. Still, this does not mean, though, that the national agenda of economic integration or social development was not pursued. During the same time rural credit offered to local agricultural producers in Amazonian states, it went up from 500 million reais a year in 1999 to over 4 billion a year by 2012 [@capobianco2021]. Now-a-days, it has reached unprecedented levels.

@@

This paper answers how the Amazon has been constructed as a problem in Brazilian presidential speeches since 1985. Our findings show that these constructions are not monolithic within governments, in time, or across different locations. Conceptually, this contributes to understanding the social construction of the Amazon in discourses and how these might relate to outcomes such as policies and environmental outcomes. Empirically, we build the first comprehensive overview of the Amazon in presidential discourse.

@@

# stargazer::stargazer(model, type = "text",
#                      covariate.labels = c("Brasilia", "International", "Non-AM state",
#                                           "Election", "Deforestation", "Inflation"),
#                      column.labels = c("EI", "SD", "Sov", "EI-Con", "EI-SD", "Sov-Con", "SD-EI-Con",
#                                        "Sov-EI", "SD-Con", "Other"),
#                      dep.var.labels.include = FALSE)

@@

  kable_styling(latex_options = c("striped", "scale_down"),
                position= "center", font_size = 9, full_width = TRUE) %>%
  column_spec(1, width="3 cm", bold=TRUE, latex_valign = "m") %>%
  column_spec(2, width="2 cm", latex_valign = "m") %>%
  column_spec(3, width = "2 cm", latex_valign = "m") %>%
  column_spec(4, width = "2 cm", latex_valign = "m") %>%
  column_spec(5, width="2 cm", latex_valign = "m") %>%
  column_spec(6, width = "2 cm", latex_valign = "m") %>%
  column_spec(7, width = "2 cm", latex_valign = "m") %>% 
  column_spec(8, width="2 cm", latex_valign = "m") %>%
  column_spec(9, width = "2 cm", latex_valign = "m") %>%
  column_spec(10, width = "2 cm", latex_valign = "m") %>% 
  column_spec(11, width = "2 cm", latex_valign = "m")
